Plant-pollintor networks worldwide are composed by the same specific building

### 2021-09-29 Draft/Draft.Rmd
Results of figure 4 commented

@@

@@

Traditionally, plant-pollinator research has focused on direct interactions but overlooked indirect interactions (the mediated effect between two species by a third species), such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g. interaction frequency) or in single topological indices (e.g. nestedness). Nonetheless, the emerging framework of network motifs in plant-pollinator research, the building blocks of a network that depict subsets of interactions [@milo2002], allows to consider both direct and indirect interactions [@simmons2019]. The analysis of motifs (meso-scale) in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in plant-pollinator networks (over- and under- represented motifs) are still unknown.

@@

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks. Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) over and under-represented motifs; (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

@@

First, plant species were grouped thorugh hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. We used the trait dataset collated in Lanuza et al., (unpublished) that comprised 1,506 plant species and contained the same 60 plant-pollinator networks used in this study plus some non-weighted and weighted metawebs (see **Table S1** Lanuza et al., unpublished). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30%; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. Previous to the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera, (vi) Coleoptera, (vii). However, a minor set of species belonged to other functional groups that were considered in the analysis as lizards, birds and other insects but not discussed further because of their low representation in the set of analysed networks.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE, fig.cap="**Figure 1.** Adapted figure of Simmons et al., 2019 with all the possible motifs from two to five species in bipartite networks. There is a total of 17 possible motifs with 46 different positions denoted within each node.", fig.width=10,fig.height=8}
library(tidyverse)
library(RColorBrewer)
source("../Scripts/Scripts_Alfonso/plot_motif_positionsV2.R")

@@

To test the effect of singletones on the motif frequencies and motif functional groups' positions, we applied the previous analyses to the same set of networks (**Table S1**) but considering only interactions whose frequency was greater than one (64.98% of interactions; see **Figures S3 and S4**).

@@

In total, there were 1126 species of floral visitor with 6325 interactions recorded with plants. From highest to lowest records of interactions of floral visitors with plants we found: bees with 2256 interactions, non-syrphid Diptera with 1768 interactions, syrphids with 845 interactions, Lepidoptera with 437 interactions, Coleoptera with 432 interactions and non-bee Hymenoptera with 362 interactions.

@@

Most motifs were under- and over-represented (close to the 1st and 99th percentile, respectively) in the comparison between empirical and simulated networks (**Figure 2**). Motifs 3, 5, 9, 10 and 14 were under-represented in empirical networks, that is, all were close to the 1st percentile and under the 25th percentile. Remarkably, four out of five of these motifs belonged to the largest path length classification (i.e., core-peripheral). In addition, motifs 2, 6, 7, 16 and 17 were over-represented, all over the 75th percentile and thus the closest to the 99th percentile. In contrast to the under-represented motifs, over-represented motifs belonged to the two shortest path length groups (i.e., complete and fan). The remaining motifs (i.e., 4, 8, 11, 12, 13, and 15) were between the 25th and 75th percentile but closest to over-representation. Remarkably, the exclusion of singletones from this analysis did not change the observed results (**Figure S1**).

@@

p <- ggplot()+geom_point(data = d_2, aes(y=factor(motif), 100*percentil_sizeclass), shape=21,color="black",fill="azure2",size=1.75, alpha=1)+
  geom_errorbar(data = d_2,aes(y = factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=0.8,size=0.8,color="black") +theme_bw()+
  geom_point(data = d_2,aes(y = as.factor(motif), x=100*mean), 
             size=2,color="black")+labs(y="Two to five node motifs", x = "Percentile")+
  theme( strip.background = element_blank())+
  theme(strip.text = element_text(face="bold", size=10, hjust=0),
        panel.border=element_rect(color="black",size=1.2))

@@

Although motifs positions determined over- and under- representation independently of the identity of functional groups for most cases (rows in **Figure 3**), we found differences in the average percentile within floral visitors and pollinators functional groups (columns in **Figure 3**). These differences were more marked for floral visitors functional groups. The group of Bees was the most over-represented, followed by non-syrphids Diptera, syrphids, Coleoptera, non-bee hymenoptera and finally Lepidoptera which was the most under-represent functional group on the different motif positions. Although for plants the differences across groups were less pronounced, from most represented to lowest we found self incompatible perennials with large flowers, small outcrossing perennials, tall plants with unisexual flowers, selfing herbs and short lived outcrossers with long zygomorphic flowers.

@@

@@

  # Define the number of colors you want
  nb.cols <- 5
  mycolors <- c(brewer.pal(6, "Paired"),brewer.pal(5, "Dark2"))

  ggplot() +
    scale_shape_identity() +
    geom_segment(data = link_df, aes(x=xdown, y=ydown, xend = xup, yend = yup), 
                 size = 1, color = "grey20")+  geom_scatterpie(aes(x=x, y=y, group=position,r=.45), 
    data = d_nodes_pollinator, cols=as.character(c("Bee","Coleoptera","Lepidoptera","Non-bee-Hymenoptera", "Non-syrphids-diptera","Syrphids"))) +
    geom_scatterpie(aes(x=x, y=y, group=position,r=0.45), 
                    data = d_nodes_plant,
                    cols=as.character(c("1","2","3","4","5")))+
    coord_fixed()+
    scale_fill_manual(values = mycolors,"test",
      breaks=c("Bee","Coleoptera","Lepidoptera",
               "Non-bee-Hymenoptera", "Non-syrphids-diptera",
               "Syrphids","1","2","3","4","5"), labels=c("Bee","Coleoptera","Lepidoptera",
"Non-bee-Hymenoptera", "Non-syrphids-diptera","Syrphids", 
          "Selfing herbs","Small outcrossing perennials","Self-incomp. perennials\nwith large flowers",
          "Tall plants with small\nunisexual flowers","Short-lived outcrossers with\nlong zygomorphic flowers"))+
    xlim(-1.5,4)+ ylim(-0.5,2.5)+
    theme(panel.background = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank())+
    labs(x=NULL,y=NULL)+
    guides(color = FALSE,fill=guide_legend(title="Functional groups"))+
    ggtitle(paste0("Motif ",motif_number))+
    theme(plot.title = element_text(hjust = 0.5),legend.key.size = unit(0.5, 'cm'))+
    theme(plot.margin=unit(c(0,0,0,0),"cm"))
}

@@

  # Define the number of colors you want
  mycolors <- c(brewer.pal(6, "Paired"),brewer.pal(5, "Dark2"))

@@

>>>>

__Jose B. Lanuza^1,2^__  | __Alfonso Allen-Perkins^2,3^__ |  __Romina Rader^1^__ | __Ignasi Bartomeus^2^__

1 School of Environmental and Rural Science, University of New England, Armidale, New South Wales 2350, Australia.

2 Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain.

3 Departamento de Ingeniería Eléctrica, Electrónica, Automática y Física Aplicada, ETSIDI, Universidad Politécnica de Madrid, 28040 Madrid, Spain.

@@

\doublespacing

@@

Traditionally, plant-pollinator research has focused on direct interactions but overlooked indirect interactions (i.e., the mediated effect between two species by a third species), such as facilitative or competitive interactions between plants for pollinators [@moeller2004; @sargent2008]. Despite the widespread nature of indirect interactions in ecological communities [@strauss1991], plant-pollinator research often fails to finely capture those indirect interactions with the conventional analytical tools that condense the information either by species (e.g., interaction frequency) or in single topological indices (e.g., nestedness). Nonetheless, the emerging framework of network motifs in plant-pollinator research, the building blocks of a network that depict subsets of interactions [@milo2002], allows to consider both direct and indirect interactions [@simmons2019]. The analysis of motifs (meso-scale) in plant-pollinator networks have revealed that the different ecological processes that govern species interactions (e.g., species abundances versus trait-matching) can lead to different patterns of indirect interactions [@simmons2020]. Yet, the global patterns of indirect interactions in plant-pollinator networks (over- and under- represented motifs) are still unknown.

@@

Here, we used 60 plant-pollinator networks from 18 different studies and 14 countries, alongside functional groups of plants and floral visitors based on functional traits and taxonomic rank, respectively. To obtain plant functional groups, we used a comprehensive dataset that included floral, reproductive and vegetative traits compiled on a larger set of networks. Floral visitor functional groups were divided by the main taxonomical groups that differed in life form and behaviour. Once we split the different plant-pollinator networks into their motif elements, we explored: (i) the overall motif patterns (over and under-represented motifs); (ii) over and under-represented functional groups in the different motif positions; (iii) the probability of finding plant and floral visitor functional groups in the different motif positions; and, (iv) over- and under- represented functional group combinations within the different motifs.

---
# THIS IS A COMMENT
# Each os this points (i to iv) corresponds to a figure
# I think that ideally we can have that structure on methods and results (excluding functional groups)
# so far we have (i), (ii) and (iv) on methods
# we lack an explicit mention to (iii) 
# maybe merge it with another research question and just have 3?
# What do you think? 
---

@@

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. We used the trait dataset collated in Lanuza et al., (unpublished) that comprised 1,506 plant species and contained the same 60 weighted plant-pollinator networks used in this study plus 4 additional large unweighted metawebs (see **Table S1** Lanuza et al., unpublished). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30%; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. Previous to the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) Hymenoptera-Anthophila (bees), (ii) Hymenoptera-non-Anthophila (other non-bee Hymenoptera), (iii) Syrphidae-Diptera, (iv) non-Syrphidae-Diptera, (v) Lepidoptera and (vi) Coleoptera. However, a minor set of species belonged to other functional groups that were considered in analyses but not discussed further because of their low representation (3.55% of the total interactions recorded in the full set of networks). These functional groups were 'lizards', 'birds' and 'other insects'. This last group was formed by a mix of uncommon insect taxa on the full set of networks.

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, fig.pos="t", warning=FALSE,out.width="75%",fig.align="center", fig.cap="\\textbf{Figure 1.} Adapted figure of Simmons et al., 2019 with all the possible motifs from two to five species in bipartite networks. There is a total of 17 possible motifs with 46 different positions denoted within each node.", fig.width=10,fig.height=8}
library(tidyverse)
library(RColorBrewer)
source("../Scripts/Scripts_Alfonso/plot_motif_positionsV2.R")

@@

To test the effect of singletones on the motif frequencies and motif functional groups' positions, we applied the previous analyses to the same set of networks (**Table S1**) but considering only interactions whose frequency was greater than one (64.98% of interactions; see **Figures S1 and S2**).

@@

In total, there were 1126 species of floral visitor with 6325 interactions recorded with plants. There were 2256 bee interactions with plants, 1768 non-syrphid Diptera interactions, 845  syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee Hymenoptera interactions.

@@

Most motifs were under- and over-represented (close to the 1st and 99th percentile, respectively) in the comparison between empirical and simulated networks (**Figure 2**). Motifs 3, 5, 9, 10 and 14 were under-represented in empirical networks, that is, all were close to the 1st percentile and under the 25th percentile. Further, four out of five of these motifs belonged to the largest path length classification (i.e., core-peripheral). In addition, motifs 2, 6, 7, 16 and 17 were over-represented, all over the 75th percentile and thus the closest to the 99th percentile. In contrast to the under-represented motifs, over-represented motifs belonged to the two shortest path length groups (i.e., complete and fan). The remaining motifs (i.e., 4, 8, 11, 12, 13, and 15) were between the 25th and 75th percentile but closest to over-representation. Notably, the exclusion of singletones from this analysis did not change the observed results (**Figure S1**).

@@

p <- ggplot()+geom_point(data = d_2, aes(y=factor(motif), 100*percentil_sizeclass), shape=21,color="black",fill="azure2",size=1.75, alpha=0.3)+
  geom_errorbar(data = d_2,aes(y = factor(motif), xmin=100*(mean-SE), xmax=100*(mean+SE)), 
                width=0.8,size=0.8,color="black") +theme_bw()+
  geom_point(data = d_2,aes(y = as.factor(motif), x=100*mean), 
             size=2,color="black")+labs(y="Two to five node motifs", x = "Percentile")+
  theme( strip.background = element_blank())+
  theme(strip.text = element_text(face="bold", size=10, hjust=0),
        panel.border=element_rect(color="black",size=1.2))

@@

Although motifs positions determined over- and under- representation independently of the identity of functional groups for most cases (rows in **Figure 3**), we found differences in the average percentile within floral visitors and pollinators functional groups (columns in **Figure 3**). These differences were more marked for floral visitors functional groups. The group of Bees was the most over-represented, followed by non-syrphids Diptera, syrphids, Coleoptera, non-bee hymenoptera and finally Lepidoptera which was the most under-represent functional group on the different motif positions. Although for plants the differences across groups were less pronounced, from most represented functional group on the different motif positions to lowest we found self incompatible perennials with large flowers, small outcrossing perennials, tall plants with unisexual flowers, selfing herbs and short lived outcrossers with long zygomorphic flowers.

@@

The different observed frequencies on the different motif positions showed a hirarchical order of probabilities for floral visitors (**Figure 4**). That is, the functional group of 'bees' was the most abundant one on all motif positions followed by non-syrphid Diptera which was after bees, the second most frequent functional group on all positions. Then, the group of syrphids was the most abundant in all positions but two (positions 28 and 31 from motifs 11 and 12, respectively). After syrphids, Coleoptera was the most frequent group on the different positions followed by non-bee Hymenoptera and finally, Lepidoptera which was the less frequent group on all positions but two (positions 12 and 22 from motifs 5 and 9, respectively). Remarkably, plant functional groups were more variable on the different motif positions. However, there were also predominant functional groups found across the different motif positions. The three most frequent groups were 'tall plants with small unisexual flowers', 'self-incompatible perennials with large flowers' and 'small outcrossing perennials'. On the contrary, the groups of 'selfing herbs' and 'short lived outcxrossers with long zygomorphic flowers' had the lowest probability to be present on the different motifs plant positions.

@@

  # Define the number of colors you want
 mycolors <- c("#A6CEE3", "chocolate2",  "gold2", "#1F78B4","#B2DF8A",
                     "gray58", "lightgoldenrod2","thistle2", "burlywood","lightcoral", "olivedrab3")

  ggplot() +
    scale_shape_identity() +
    geom_segment(data = link_df, aes(x=xdown, y=ydown, xend = xup, yend = yup), 
                 size = 1, color = "grey20")+  geom_scatterpie(aes(x=x, y=y, group=position,r=.45), 
    data = d_nodes_pollinator, cols=as.character(c("Bee","Coleoptera","Lepidoptera","Non-bee-Hymenoptera", "Non-syrphids-diptera","Syrphids"))) +
    geom_scatterpie(aes(x=x, y=y, group=position,r=0.45), 
                    data = d_nodes_plant,
                    cols=as.character(c("1","2","3","4","5")))+
    coord_fixed()+
    scale_fill_manual(values = mycolors,"test",
      breaks=c("Bee","Coleoptera","Lepidoptera",
               "Non-bee-Hymenoptera", "Non-syrphids-diptera",
               "Syrphids","1","2","3","4","5"), labels=c("Bee","Coleoptera","Lepidoptera",
"Non-bee-Hymenoptera", "Non-syrphids-diptera","Syrphids", 
          "Selfing herbs","Small outcrossing perennials","Self-incomp. perennials\nwith large flowers",
          "Tall plants with small\nunisexual flowers","Short-lived outcrossers with\nlong zygomorphic flowers"))+
    xlim(-1.5,4)+ ylim(-0.5,2.5)+
    theme(panel.background = element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.title = element_text(size=12, face="bold"),legend.background = element_rect(size=0.5, linetype="solid",colour ="black"))+
    labs(x=NULL,y=NULL)+
    guides(color = FALSE,fill=guide_legend(title="Functional groups"))+
    ggtitle(paste0("Motif ",motif_number))+
    theme(plot.title = element_text(hjust = 0.5),legend.key.size = unit(0.5, 'cm'))+
    theme(plot.margin=unit(c(0,0,0,0),"cm"))
}

@@

  # Define the number of colors you want
 mycolors <- c("#A6CEE3", "chocolate2",  "gold2", "#1F78B4","#B2DF8A",
                     "gray58", "lightgoldenrod2","thistle2", "burlywood","lightcoral", "olivedrab3")

@@

##DISCUSSION
